---
title: "Monthly pm2.5 in Bangkok : Comparing means"
output: html_notebook
---

[Inspired by prasertcbs](https://www.youtube.com/watch?v=-MiD-EWfdiI&list=PLoTScYm9O0GGat89RT9NMjW7sqFz84XSk&index=4&t=317s)  
[Raw data here](http://berkeleyearth.lbl.gov/air-quality/maps/cities/Thailand/Bangkok)

```{r setup , include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(tidyverse)
library(lubridate)
library(fpp2)
library(ggpubr)
library(rstatix)
```

### From this plot
```{r , echo=FALSE}
pm2_5 %>%
  mutate(month = as.factor(month(local_date_time))) %>%
  ggplot(aes(pm2.5,month)) +
  geom_density_ridges(aes(fill = month),quantile_lines = TRUE, quantiles = 0.5) +
  theme(legend.position = "none") +
  labs(title = "Bangkok PM 2.5 : Monthly Distribution " , caption = "data : berkeleyearth.org")
```

## Question : Does monthly seasonal variation in pm 2.5 exist ? 

### Descriptive Stat
```{r , echo=FALSE}
pm2_5.monthly %>%
  group_by(month) %>%
  get_summary_stats(pm2.5 , type = "mean_sd")
```

## Step 1 : Distribution is Normal or not ? 
We can check this in 2 ways : 1) Inspect QQ plot , 2) Perform Shapiro's test

### 1) QQ plot
```{r qq_plot }
ggqqplot(pm2_5.monthly , "pm2.5", facet.by = "month")
```

Data not in standard error range , It looks like non-normal distribution 

### 2) Shapiro's test 
$H_0$ : Normal distribution in sample

```{r shapiro_test}
pm2_5.monthly %>%
  group_by(month) %>%
  shapiro_test(pm2.5) %>%
  mutate(normal_dist = if_else(p > 0.05, "normal","non-normal" ))
```
**All non-normal distribution , comfirmed !! **   

Thus, We have to use non-parametric alternative to the one-way ANOVA test,
**Kruskal-Wallis Test.**

# Kruskal-Wallis Test
[Data Novia](https://www.datanovia.com/en/lessons/kruskal-wallis-test-in-r/)

$$H_0 : \mu_1 = \mu_2 = \mu_3 = \ ...\  = \mu_{12}$$
$$H_0 : \mu_1 \neq \mu_2 \neq \mu_3 \neq \ ...\  \neq \mu_{12}$$

```{r }
pm2_5.monthly %>% rstatix::kruskal_test(pm2.5 ~ month)
```
**P-value very close to 0 ?**

**Use Base R instead !**
```{r}
kruskal.test(pm2.5 ~ month, pm2_5.monthly)
```
**P-value is significant. Thus,there are at least one difference in mean pm2.5 between months.**  
**But, how much of the differences ? **

### Kruskal effectsize calculation
```{r}
pm2_5.monthly %>% rstatix::kruskal_effsize(pm2.5 ~ month)
```
**Meaning : 44% of variation in mean pm2.5 can be predicted by month. This effect size is large !!**

## Multiple pairwise-comparisons

From the output of the Kruskal-Wallis test, we know that there is a significant difference between groups, but we don’t know which pairs of groups are different.

### Dunn's test
```{r}
dunn_pm2_5.month <- pm2_5.monthly %>% 
        rstatix::dunn_test(pm2.5 ~ month, p.adjust.method = "bonferroni")
```
```{r}
dunn_pm2_5.month %>%
  mutate(mean_diff = if_else(p.adj > 0.001, "Non-Significant","Significant" )) %>%
  mutate(across(starts_with("g"),as_factor)) %>%
  select(".y.",starts_with("group"),p.adj,mean_diff) %>%
  arrange(desc(p.adj)) 

```
### Visualize significant Dunn's test
```{r}
dunn_pm2_5.month %>%
  mutate(mean_diff = if_else(p.adj > 0.001, "Non-Significant","Significant" )) %>%
  mutate(across(starts_with("g"),as_factor)) %>%
  ggplot(aes(group1,group2,color = mean_diff, size = p.adj)) + 
  geom_point() + scale_size(name = NULL , breaks = NULL) +
  xlab("month") + ylab("month") + 
  labs(title = "Dunn's test comparison between means of pm2.5", caption = "Significant at P-value < 0.001")
```
### Final Report
* There was a statistically significant differences of pm2.5 mean value between months as assessed using the Kruskal-Wallis test (p < 0.001). Pairwise Dunn's test between months showed that all the difference between months was significant (Dunn’s test, p < 0.001) except for 7 pairs of months as indicated by red dot in above graph. 

> Light Bridge : R-Studio 1.3.959 (mac)
